---
title: "Final_Project"
author: "Lisa McDaniel"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message=FALSE}

library(readxl)
library(tidyverse)
library(dplyr)

FiveYear <- read_excel('./2021_2025_5YrProg_Highway_Program.xlsx')

str(FiveYear) 
names(FiveYear)

FiveYear <- FiveYear %>% 
  fill(Location, .direction = "down") %>% 
  rename(District = `Iowa DOT District`, Route = `Project Route`, 
         Work = `Type of Work`) %>%
  group_by(County, District, Route, Location) %>%
  fill(Miles, .direction = "down") %>%
  mutate(District = as.numeric(District))
  

```


## Allan Juarez

<font size="3"> 

* Type, number, Iowa DOT District, County, route, location and estimated costs of projects planned between 2021 – 2025.  Where is the money going and how much?

```{r }

```

## John King

<font size="3"> 

* What work types are associated with interchanges?  Costs? What routes are interchanges being proposed on?

```{r }

```

## Adam Brandt

<font size="3"> 

* Projects < $50 million, projects > $50 million & < $100 million, major projects > $100.  What is the most expensive project in each District?  The least expensive?

```{r }

```

## Colin Willenborg(Project Coordinator)

<font size="3"> 

* Project types and costs between Iowa DOT Districts and years.

```{r}

```

## Lisa McDaniel

<font size="5"> Bridge Assets

<font size="3"> 

* How/do bridge/culvert work types relate to the federal work types:  construction, replacement, reconstruction, maintenance or preservation?  

* How do funding levels in the 5 Year Program compare to those in the Asset Management Plan?

* Compare the five year total of the 5 Year Program to the Asset Management Plan

```{r }

```

<font size="5"> Pavement Assets

<font size="3"> 

* How/do pavement work types relate to the federal work types:  construction, replacement, reconstruction, maintenance or preservation?  

* How do funding levels in the 5 Year Program compare to those in the Asset Management Plan?

* Compare the five year total of the 5 Year Program to the Asset Management Plan

```{r }

```