---
title: "Final_Project"
author: "Lisa McDaniel"
date: "11/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r message=FALSE}

library(readxl)
library(tidyverse)
library(dplyr)

FiveYear <- read_excel('./2021_2025_5YrProg_Highway_Program.xlsx')

str(FiveYear) 
names(FiveYear)

FiveYear <- FiveYear %>% 
  fill(Location, .direction = "down") %>% 
  rename(District = `Iowa DOT District`, Route = `Project Route`, 
         Work = `Type of Work`) %>%
  group_by(County, District, Route, Location) %>%
  fill(Miles, .direction = "down") %>%
  mutate(District = as.numeric(District))
  

```


## Allan Juarez

<font size="3"> 

* Type, number, Iowa DOT District, County, route, location and estimated costs of projects planned between 2021 – 2025.  Where is the money going and how much?

```{r }
states <- map_data('state')
us_counties <- map_data("county")

twentyone <- FiveYear %>% 
  select(County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2021") %>%
  rename(year ="2021")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(subregion) %>% 
  summarize(total = sum(year,na.rm=TRUE))

 
iowa <- c("iowa")
us_counties <- us_counties %>% filter(region %in% iowa)
twentyone <- inner_join(twentyone, us_counties, by = "subregion")
 twentyone %>% ggplot(data = twentyone,
            mapping = aes(x = long, y = lat,
                          group = group, fill = total))+ geom_polygon(color = "gray90", size = 0.1) 
 
 twentytwo <- FiveYear %>% 
  select(County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2022") %>%
  rename(year ="2022")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(subregion) %>% 
  summarize(total = sum(year,na.rm=TRUE))

 
iowa <- c("iowa")
us_counties <- us_counties %>% filter(region %in% iowa)
twentytwo <- inner_join(twentytwo, us_counties, by = "subregion")
 twentytwo %>% ggplot(data = twentytwo,
            mapping = aes(x = long, y = lat,
                          group = group, fill = total))+ geom_polygon(color = "gray90", size = 0.1)
 
twentythree <- FiveYear %>% 
  select(County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2023") %>%
  rename(year ="2023")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(subregion) %>% 
  summarize(total = sum(year,na.rm=TRUE))

 
iowa <- c("iowa")
us_counties <- us_counties %>% filter(region %in% iowa)
twentythree <- inner_join(twentythree, us_counties, by = "subregion")
 twentythree %>% ggplot(data = twentythree,
            mapping = aes(x = long, y = lat,
                          group = group, fill = total))+ geom_polygon(color = "gray90", size = 0.1) 
 
twentyfour <- FiveYear %>% 
  select(County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2024") %>%
  rename(year ="2024")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(subregion) %>% 
  summarize(total = sum(year,na.rm=TRUE))

 
iowa <- c("iowa")
us_counties <- us_counties %>% filter(region %in% iowa)
twentyfour <- inner_join(twentyfour, us_counties, by = "subregion")
 twentyfour %>% ggplot(data = twentyfour,
            mapping = aes(x = long, y = lat,
                          group = group, fill = total))+ geom_polygon(color = "gray90", size = 0.1)
 
 twentyfive <- FiveYear %>% 
  select(County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2025") %>%
  rename(year ="2025")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(subregion) %>% 
  summarize(total = sum(year,na.rm=TRUE))

 
iowa <- c("iowa")
us_counties <- us_counties %>% filter(region %in% iowa)
twentyfive <- inner_join(twentyfive, us_counties, by = "subregion")
 twentyfive %>% ggplot(data = twentyfive,
            mapping = aes(x = long, y = lat,
                          group = group, fill = total))+ geom_polygon(color = "gray90", size = 0.1)

```

## John King

<font size="3"> 

* What work types are associated with interchanges?  Costs? What routes are interchanges being proposed on?

```{r }

```

## Adam Brandt

<font size="3"> 

* Projects < $50 million, projects > $50 million & < $100 million, major projects > $100.  What is the most expensive project in each District?  The least expensive?

```{r }

```

## Colin Willenborg(Project Coordinator)

<font size="3"> 

* Project types and costs between Iowa DOT Districts and years.

```{r}
head(FiveYear)
districts <- FiveYear %>% 
  select(County, District) %>% 
  mutate(subregion = tolower(County))

county_map <- map_data("county")
ia <- subset(county_map, region =="iowa")

districts_2021 <- FiveYear %>% 
  select(District, County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2021") %>%
  rename(year ="2021")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(District) %>% 
  summarize(total = sum(year,na.rm=TRUE))

ia <- inner_join(ia, districts, by="subregion")

districts_2021 <- inner_join(districts_2021, ia, by = "District")

districts_2021 %>% ggplot(aes(x = long, y = lat, fill=District)) + geom_polygon(aes(group=group,fill = total), color="gray60", size=0.1)

districts_2022 <- FiveYear %>% 
  select(District, County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2022") %>%
  rename(year ="2022")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(District) %>% 
  summarize(total = sum(year,na.rm=TRUE))

districts_2022 <- inner_join(districts_2022, ia, by = "District")

districts_2022 %>% ggplot(aes(x = long, y = lat, fill=District)) + geom_polygon(aes(group=group,fill = total), color="gray60", size=0.1)

# 2023
districts_2023 <- FiveYear %>% 
  select(District, County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2023") %>%
  rename(year ="2023")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(District) %>% 
  summarize(total = sum(year,na.rm=TRUE))

districts_2023 <- inner_join(districts_2023, ia, by = "District")

districts_2023 %>% ggplot(aes(x = long, y = lat, fill=District)) + geom_polygon(aes(group=group,fill = total), color="gray60", size=0.1)

# 2024
districts_2024 <- FiveYear %>% 
  select(District, County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2024") %>%
  rename(year ="2024")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(District) %>% 
  summarize(total = sum(year,na.rm=TRUE))

districts_2024 <- inner_join(districts_2024, ia, by = "District")

districts_2024 %>% ggplot(aes(x = long, y = lat, fill=District)) + geom_polygon(aes(group=group,fill = total), color="gray60", size=0.1)

# 2025
districts_2025 <- FiveYear %>% 
  select(District, County,'Latitude\r\n(Decimal(', "Longitude\r\n(Decimal)","2025") %>%
  rename(year ="2025")%>%
  mutate(subregion = tolower(County)) %>% 
  group_by(District) %>% 
  summarize(total = sum(year,na.rm=TRUE))

districts_2025 <- inner_join(districts_2025, ia, by = "District")

districts_2025 %>% ggplot(aes(x = long, y = lat, fill=District)) + geom_polygon(aes(group=group,fill = total), color="gray60", size=0.1)

```

## Lisa McDaniel

<font size="5"> Bridge Assets

<font size="3"> 

* How/do bridge/culvert work types relate to the federal work types:  construction, replacement, reconstruction, maintenance or preservation?  

* How do funding levels in the 5 Year Program compare to those in the Asset Management Plan?

* Compare the five year total of the 5 Year Program to the Asset Management Plan

```{r }

```

<font size="5"> Pavement Assets

<font size="3"> 

* How/do pavement work types relate to the federal work types:  construction, replacement, reconstruction, maintenance or preservation?  

* How do funding levels in the 5 Year Program compare to those in the Asset Management Plan?

* Compare the five year total of the 5 Year Program to the Asset Management Plan

```{r }

```