---
title: "DataWork"
author: "John King"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r, warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(maps)
library(mapproj)

yearplan <- read_excel('./2021_2025_5YrProg_Highway_Program.xlsx')


cleanplan <- yearplan %>%
  fill(Location, .direction = "down") %>% 
  rename(District = `Iowa DOT District`, Route = `Project Route`, 
         Work = `Type of Work`) %>%
  group_by(County, District, Route, Location) %>%
  fill(Miles, .direction = "down") %>%
  mutate(District = as.numeric(District)) %>%
  select(State, County, District, Route, Location, 
         Miles, Work, `2021`, `2022`, `2023`, `2024`, `2025`)
cleanplan['totalProjection'] <- rowSums(cleanplan[, c("2021", "2022", "2023", "2024",
                                                      "2025")], na.rm = TRUE)
countyLines <- map_data("county")


#Cleaning City Population data
cities <- read_excel("./SUB-IP-EST2019-ANNRES-19.xlsx")
cities <- cities %>% select(`table with row headers in column A and column headers in rows 3 through 4 (leading dots indicate sub-parts)`, `...13`) %>% 
  rename(Population = "...13", Area = "table with row headers in column A and column headers in rows 3 through 4 (leading dots indicate sub-parts)")
cities <- cities[-c(1,2,3),]
cities <- cities %>% separate(Area, into = c("City", "State"), sep = ", ") %>% select(-State)
```

## Work Types Associated With Interchanges

```{r, warning=FALSE, message=FALSE}

  
interchange <- cleanplan %>% filter(grepl('INTERCHANGE', Location))

interchange["County"] <- tolower(interchange$County)
interchange

interchange %>% group_by(Work) %>% summarise(sum = n()) %>% 
  ggplot(aes(x=Work, weight=sum)) + geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))

interchange %>% group_by(Location) %>% summarise(WorkSum = n()) %>% 
  filter(WorkSum != 1)


# Is this useful? Maybe for report...
interchange %>% group_by(Location) %>% summarise(WorkSum = n()) %>% 
  filter(WorkSum != 1) %>% 
  ggplot(aes(x=Location, weight = WorkSum)) + geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) + 
  scale_x_discrete(label=abbreviate)
```

Bridge Deck Overlays seem to be the most common type of work on the interchanges. It seems that a great number of planned projects are related to maintenance or additions for the sake of drivers (traffic signs, traffic signals, lighting).

## Counties With Proposed Interchanges

```{r, warning=FALSE, message=FALSE}

iowaLines <- countyLines %>% filter(region == "iowa")
iowaLines <- iowaLines %>% select(long, lat, group, order, subregion)
joinedInter <- iowaLines %>% left_join(interchange, by = c(subregion = "County"))

joinedInter["present"] <- ifelse(is.na(joinedInter$District), "no", "yes")

cityData <- map_data("us.cities")

joinedInter %>% ggplot(aes(x = long, y = lat, fill = present)) + 
  geom_polygon(aes(group = group)) + coord_map()

```

## Cost of Interchanges

```{r, warning=FALSE, message=FALSE}

interchange %>% ggplot(aes(y = totalProjection, x= Location)) + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1))
```

## Routes with proposed Interchanges


