---
title: "DataWork"
author: "John King"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

```{r, warning=FALSE, message=FALSE}
library(readxl)
library(tidyverse)
library(osmdata)
library(sf)
library(maps)
library(mapproj)
#library(ggmap)
#library(ggthemes)

yearplan <- read_excel('./2021_2025_5YrProg_Highway_Program.xlsx')


cleanplan <- yearplan %>%
  fill(Location, .direction = "down") %>% 
  rename(District = `Iowa DOT District`, Route = `Project Route`, 
         Work = `Type of Work`) %>%
  group_by(County, District, Route, Location) %>%
  fill(Miles, .direction = "down") %>%
  mutate(District = as.numeric(District)) %>%
  select(State, County, District, Route, Location, 
         Miles, Work, `2021`, `2022`, `2023`, `2024`, `2025`)
cleanplan['totalProjection'] <- rowSums(cleanplan[, c("2021", "2022", "2023", "2024",
                                                      "2025")], na.rm = TRUE)
countyLines <- map_data("county")


#Cleaning City Population data
cities <- read_excel("./SUB-IP-EST2019-ANNRES-19.xlsx")
cities <- cities %>% select(`table with row headers in column A and column headers in rows 3 through 4 (leading dots indicate sub-parts)`, `...13`) %>% 
  rename(Population = "...13", Area = "table with row headers in column A and column headers in rows 3 through 4 (leading dots indicate sub-parts)")
cities <- cities[-c(1,2,3),]
cities <- cities %>% separate(Area, into = c("City", "State"), sep = ", ") %>% select(-State)
```

## Work Types Associated With Interchanges

```{r, warning=FALSE, message=FALSE}

  
interchange <- cleanplan %>% filter(grepl('INTERCHANGE', Location))

interchange["County"] <- tolower(interchange$County)
interchange["Route"] <- as.character(interchange$Route)
interchange

interchange %>% group_by(Work) %>% summarise(sum = n()) %>% 
  ggplot(aes(x=Work, weight=sum)) + geom_bar() + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(title = "Count of Interchange Work Plans per Work Type", x = "Type of Work")

interchange %>% group_by(Location) %>% summarise(WorkSum = n()) %>% 
  filter(WorkSum != 1)


# Is this useful? Maybe for report...
interchange %>% group_by(Location) %>% summarise(WorkSum = n()) %>% 
  filter(WorkSum != 1) %>% 
  ggplot(aes(x=Location, weight = WorkSum)) + geom_bar() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) + 
  scale_x_discrete(label=abbreviate)
```

Bridge Deck Overlays seem to be the most common type of work on the interchanges. It seems that a great number of planned projects are related to maintenance or additions for the sake of drivers (traffic signs, traffic signals, lighting).

## Counties With Proposed Interchanges

```{r, warning=FALSE, message=FALSE}

iowaLines <- countyLines %>% filter(region == "iowa")
iowaLines <- iowaLines %>% select(long, lat, group, order, subregion)
joinedInter <- iowaLines %>% left_join(interchange, by = c(subregion = "County"))

joinedInter["presentInIntersectionData"] <- ifelse(is.na(joinedInter$District), "no", "yes")


joinedInter %>% ggplot(aes(x = long, y = lat, fill = presentInIntersectionData)) + 
  geom_polygon(aes(group = group)) + coord_map() #+ theme_map()

```

## Cost of Interchanges

```{r, warning=FALSE, message=FALSE}

plot1 <- interchange %>% ggplot(aes(y = totalProjection, x= Work)) + geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)) + 
  labs(x = "Type of Work", y = "Cost in Thousands of Dollars", title = "Cost of Projects by Type of Work")
  
plot1

plot1 + coord_cartesian(ylim = c(20, 16000))
```
First chart shows that the grade and pave and new bridge projects are the most expensive. Now we zoom in the plot ignoring the outliers to better see the other categories. From this chart we can see that most of the projects are relatively cheap in comparison, with only bridge replacement, grading, and paving projects frequently approaching the cost.


## Counties With Proposed Interchanges

```{r, warning=FALSE, message=FALSE}

iowaLines <- countyLines %>% filter(region == "iowa")
iowaLines <- iowaLines %>% select(long, lat, group, order, subregion)
joinedInter <- iowaLines %>% left_join(interchange, by = c(subregion = "County"))

joinedInter["presentInIntersectionData"] <- ifelse(is.na(joinedInter$District), "no", "yes")


joinedInter %>% ggplot(aes(x = long, y = lat, fill = presentInIntersectionData)) + 
  geom_polygon(aes(group = group)) + coord_map()

```


## Routes with proposed Interchanges

```{r, warning=FALSE, message=FALSE}

routeData <- interchange %>% group_by(Route) %>% summarise(sum = n())
routeData
routeData["Route"] <- factor(routeData$Route, levels = c("6", "9", "20", "25", "29", "30", "35", "59", 
                                                         "74", "80", "141", "235", "280", "380"))
routeData %>% ggplot(aes(x= Route, weight = sum)) + geom_bar() + 
  labs(title = "Count of Interchange Plans by Route")
```




We can see that the vast majority of interchange projects are being proposed on the major routes (80, 35, 380, and 29)